Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-08 article/article.tex
Commit 2024-06-08 09:12

In the relative bibliography different definition have been used for
these events, some of the are summarized by \citet{Gueymard2017}. Here,
we defined as CE events, the cases when the measured GHI at ground
level, exceeds the expected value under clear-sky conditions. Similarly,
we define as extreme cloud enhancement events (ECE), the cases when GHI
at ground level exceeds the Total Solar Irradiance (TSI). Although the
duration of these bursts can vary, from seconds to several minutes, here
we are constrained by the recorded data, to one-minute steps.

@@

The monitoring site is operating in the Laboratory of Atmospheric
Physics of the Aristotle University of Thessaloniki
(\(40^\circ\,38'\,\)N, \(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.). In this
study we present data from the period 13~April 1993 to 31~December 2023.
The GHI data were measured with a Kipp~\& Zonen CM-21 pyranometer.
During the study period, the pyranometer was independently calibrated
three times at the Meteorologisches Observatorium Lindenberg, DWD,
verifying that the stability of the instrument's sensitivity was better
than \(0.7\,\%\) relative to the initial calibration by the
manufacturer. For the acquisition of radiometric data, the signal of the
pyranometer was sampled at a rate of \(1\,\text{Hz}\). The mean and the
standard deviation of these samples were calculated and recorded for
every minute. The measurements were corrected for the zero offset
(``dark signal'' in volts), which was calculated by averaging all
measurements recorded for a period of \(3\,\text{h}\), before (morning)
or after (evening) the Sun reaches an elevation angle of \(-10^\circ\).
The signal was converted to irradiance using the ramped value of the
instrument's sensitivity between subsequent calibrations.

To further improve the quality of the irradiance data, a manual
screening was performed, in order to remove inconsistent and erroneous
recordings that can occur stochastically or systematically during the
long operation of the instrument. The manual screening was aided by a
radiation data quality assurance procedure, adjusted for the site, which
was based on the methods of Long and Shi~\citep{Long2006, Long2008a}.
Thus, problematic recordings have been excluded from further processing.
Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality
of the radiation measurements used in this study. To preserve an
unbiased representation of the data we applied a constraint, similar the
one used by \citet{CastillejoCuberos2020}. Where, for each valid hour of
day, there must exist at least 45 minutes of valid measurements,
including nighttime, when near sunrise and sunset. Days with less than 5
valid hours were rejected completely. Furthermore, due to the
significant measurement uncertainty, when the Sun is near the horizon,
and due to some systematic obstructions by nearby buildings, we have
excluded all measurements with solar zenith angle (SZA) greater than
\(78^\circ\).

@@

To be able to detect the CE cases, we had to establish a baseline, above
which we characterized each data point as an enhancement event, by
estimating the occurring over irradiance (OIR). The OIR, is defined as
the irradiance difference of the measured one-minute \(\text{GHI}_i\)
from the CE identification criterion
(\(\text{OIR}_i = \text{GHI}_i - \text{GHI}_\text{CSlim}\)), as defined
in Equation\nobreakspace\ref{eq:CE4}. To have an estimation, and a first
insight on the phenomenon, we experimented with two simple approaches
for the reference. The Haurwitz's model \citep{Haurwitz1945}, which is
as simple clear sky model, and we had already adjusted and had good fit
with our data \citep{Natsis2023}, and the TSI at the top of the
atmosphere. We have tested both cases by using an appropriate relative
threshold and/or an additional constant offset. The initial results,
showed that we can detect a big portion of the CE events. These results
were helpful, as they are independent from unknown factors, and helped
us to establish some criteria to further improve the CE identification.
It was evident, by inspecting the daily plot of irradiance, that changes
on the atmospheric conditions introduced numerous false positive and
false negative results. To produce a more accurate reference, we had to
take into account more factors that effect the clear sky radiation. So
we used a radiation transfer model in order to include the effects of
ADO and the water vapors.

@@

For completeness, we will describe here the main points of the radiation
simulation procedure. We used as input the spectrum of
\citet{Kurucz1994} in the range \(280\) to \(2500\,\text{nm}\), with the
Libradtran radiation transfer solver ``disort'' on a ``pseudospherical''
geometry and the ``LOWTRAN'' gas parameterization. For each combination
of conditions we use a SZA step of \(0.2^\circ\). For the atmospheric
characteristics, we iterated on combinations of AOD at
\(500\,\text{nm}\) (\(\tau_{500\text{nm}}\)) with additional offsets of
\(\pm1\) and \(\pm2\sigma\), and water column (\(w\)), also with offsets
of \(\pm1\) and \(\pm2\sigma\). We applied them on two atmospheric
profiles, from the Air Force Geophysics Laboratory (AFGL). The ``AFGL
atmospheric constituent profile, midlatitude summer'' (afglms) and the
``AFGL atmospheric constituent profile, midlatitude winter'' (afglmw)
\citep{Anderson1986}.

To create a look-up table, that aligns with our dataset, we applied some
adjustments. To account for the Sun's variability, in our one-minute GHI
measurements, we adapted each modeled value by scaling the model's input
spectrum integral, to the corresponding TSI, provided by NOAA
\citep{Coddington2005}. Also, we account for the effect of the Earth --
Sun distance on the irradiance, by using the distance calculated by the
Astropy \citep{AstropyCollaboration2022} software library. As needed, we
interpolate the resulting irradiances to the SZA of our measurements.
For each period of the year, we used the appropriate atmospheric profile
(afglms or afglmw). Finally, we calculated the clear sky irradiance
value at the horizontal plane. Thus, we were able to emulate different
atmospheric condition and levels of atmospheric clearness for the
climatological conditions of the site. With this method, the modeled
clear sky irradiances can be directly compared to each measured
one-minute value of GHI, for different conditions of atmospheric
clearness.

>>>>

In this study, we define an event as CE when the measured GHI at ground
level, exceeds the expected value under clear-sky conditions. Similarly,
we define as extreme cloud enhancement events (ECE), the cases when GHI
at ground level exceeds the Total Solar Irradiance (TSI). Although the
duration of these bursts can vary from seconds to several minutes, here
we are constrained by the temporal resolution of our data to identify
evens with duration of at minimum one-minute.

@@

The data used in this study were recorded at the monitoring site of the
Laboratory of Atmospheric Physics, Aristotle University of Thessaloniki,
in Thessaloniki, Greece (\(40^\circ\,38'\,\)N, \(22^\circ\,57'\,\)E,
\(80\,\)m~a.s.l.). The GHI data were measured with a Kipp~\& Zonen CM-21
pyranometer and cover the period 13~April 1993 to 31~December 2023.
During the study period, the pyranometer was independently calibrated
three times at the Meteorologisches Observatorium Lindenberg, DWD,
verifying that the stability of the instrument's sensitivity was better
than \(0.7\,\%\) relative to the initial calibration by the
manufacturer. For the acquisition of radiometric data, the signal of the
pyranometer was sampled at a rate of \(1\,\text{Hz}\) with the mean and
standard deviation of these samples calculated and recorded every
minute. The measurements were corrected for the zero offset (``dark
signal'' in volts), which was calculated by averaging all measurements
recorded for a period of \(3\,\text{h}\), before (morning) or after
(evening) the Sun reaches an elevation angle of \(-10^\circ\). The
signal was converted to irradiance using the ramped value of the
instrument's sensitivity between subsequent calibrations.

To further improve the quality of the irradiance data, a manual
screening was performed, in order to remove inconsistent and erroneous
recordings that can occur stochastically or systematically during the
long operation of the instrument. The manual screening was aided by a
radiation data quality assurance procedure, adjusted for the site, which
was based on the methods of Long and Shi~\citep{Long2006, Long2008a}.
Thus, problematic recordings have been excluded from further processing.
Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality
of the radiation measurements used in this study. To preserve an
unbiased representation of the data we applied a constraint, similar the
one used by \citet{CastillejoCuberos2020}. For each valid hour of day,
there must exist at least 45 minutes of valid measurements, including
nighttime, when near sunrise and sunset. Days with less than 5 valid
hours were rejected completely. Furthermore, due to the significant
measurement uncertainty in GHI when the Sun is near the horizon, and due
to some systematic obstructions by nearby buildings, we have excluded
all measurements with solar zenith angle (SZA) greater than
\(78^\circ\).

Finally, images from a sky camera have been used in the manual
inspection of the CE identification. The sky camera operates since 2012
and stores images in 5 min time steps.

@@

For the detection of CE cases we established a baseline of irradiance
above which we characterized each data point as an enhancement event and
calculated the over irradiance (OIR). The OIR is defined as the
irradiance difference of the measured one-minute \(\text{GHI}_i\) from
the CE identification criterion
(\(\text{OIR}_i = \text{GHI}_i - \text{GHI}_\text{CSlim}\)), as defined
in Equation\nobreakspace\ref{eq:CE4}. To have an estimation, and a first
insight on the phenomenon, we experimented with two simple approaches
for the reference. The Haurwitz's model \citep{Haurwitz1945}, which is
as simple clear sky model, and we had already adjusted and had good fit
with our data \citep{Natsis2023}, and the TSI at the top of the
atmosphere. We have tested both cases by using an appropriate relative
threshold and/or an additional constant offset. The initial results,
showed that we can detect a big portion of the CE events. These results
were helpful, as they are independent from unknown factors, and helped
us to establish some criteria to further improve the CE identification.
It was evident, by inspecting the daily plot of irradiance, that changes
on the atmospheric conditions introduced numerous false positive and
false negative results. To produce a more accurate reference, we had to
take into account more factors that effect the clear sky radiation. So
we used a radiation transfer model in order to include the effects of
ADO and the water vapors.

@@

For the atmospheric characteristics, we iterated on combinations of AOD
at \(500\,\text{nm}\) (\(\tau_{500\text{nm}}\)) with additional offsets
of \(\pm1\) and \(\pm2\sigma\), and water column (\(w\)), also with
offsets of \(\pm1\) and \(\pm2\sigma\). We applied them on two
atmospheric profiles, from the Air Force Geophysics Laboratory (AFGL).
The ``AFGL atmospheric constituent profile, midlatitude summer''
(afglms) and the ``AFGL atmospheric constituent profile, midlatitude
winter'' (afglmw) \citep{Anderson1986}. --\textgreater{}
